Followers Are Not Enough: A Multi-faceted Approach to Community Detection in Online Social Networks

### 2014-11-18 paper-postreview/conclusions.tex

@@

%Start of future work**
Our findings may have important implications to a common problem in social network analysis: identification of influential individuals---but further work must be carried out.
\alert{maybe move the top 10 people here}
 Many network measures of influence are based on the various types of centrality (degree, betweenness, closeness, eigenvector, etc.)~\cite{newman2009networks}. Most centralities depend explicitly on the structure of the network under consideration. 
But we have seen in our study that a structural network alone is not sufficient to capture user interaction or even activity-based influence in online social media. Thus, a na\"ive application of centrality measures to a structural network for influence detection may give rise to erroneous results. This result has been explored previously~\cite{kitsak2010identification}, and our work further highlights its importance.

>>>>

\label{sec:conc}

@@

%Start of future work**
Our findings may have important implications to a common problem in social network analysis: identification of influential individuals---but further work must be carried out.
\alert{maybe move the top 10 people here}
 Many network measures of influence are based on the various types of centrality (degree, betweenness, closeness, eigenvector, etc.)~\cite{newman2009networks}. Most centralities depend explicitly on the structure of the network under consideration. But we have seen in our study that the structural network only highlights one of many possible ways that users interact with online social media.
%But we have seen in our study that a structural network alone is not sufficient to capture user interaction or even activity-based influence in online social media.
Thus, a na\"ive application of centrality measures to a structural network may be answering a different question than the one motivating influence detection. For example, a user exhibiting high closeness centrality based on the structural network may be able to make a message visible to most other users in a social network, but this does not mean that those other users are likely to respond to or propagate that message. The importance of correctly framing network measures of influence has been explored previously~\cite{kitsak2010identification}, and our work further highlights its importance.

### 2014-11-18 paper-postreview/intro.tex

%A large body of work exists on methods for automatic detection of communities within networks~\cite{newman2004fast,newman2004finding,rosvall2008maps,blondel2008fast,LancichinettiPlos}. See~\cite{fortunato2010community} for a recent review. 
A large body of work exists on methods for automatic detection of communities within networks, {\it e.g.,} the stochastic blockmodel~\cite{Holland1983} (and its recent generalization to weighted networks~\cite{Aicher26062014}), the Newman and Girvan algorithm~\cite{newman2004finding}, clique percolation~\cite{PalEtAl05}, Infomap~\cite{Rosvall08mapsof}, Louvain~\cite{blondel2008fast}, and the recently introduced OSLOM~\cite{LancichinettiPlos}.
All these methods begin with a given network, and then attempt to uncover structure present in the network, \emph{i.e.,} they are agnostic to how the network was constructed. As opposed to this agnostic analysis, we propose---and illustrate the importance of---a \emph{multi-faceted question-focused} approach. We believe that in order to understand all of the communities present in a network, it is important to take two crucial steps: 
\begin{enumerate}
\item Form a clear question about what \emph{type} of community we wish to consider---{\it e.g.,} are we looking for groups of friends? or for groups of people having a common interest?---and then perform community detection with that question in mind.
\item Ask more than one question about the users in a network. Do several rounds of community detection, each of which is aimed at revealing a new facet of the communities present in the network. 
\end{enumerate}

@@

This is not meant to be an exhaustive list, but rather a list of some of the more common types of communities observed in social networks. We propose looking at when and how communities motivated by these different questions overlap, and whether different approaches to asking the question, ``What community are you in?'' leads to different insights about a social network. For example, a user on Twitter might connect mostly with computational social scientists, utilize the service nearly every time a particular user or group of users is active, talk mostly about machine learning, and interact solely with close friends (who may or may not be computational social scientists). Each of these different `profiles' of the user highlight different views of the user's social network, and represent different types of communities. We now consider in more detail how each of these types of questions motivates a question-specific community. Later in Section~\ref{sec:conc} we will show how asking \emph{all} of these questions---instead of just one which is the standard---gives a much deeper multi-faceted view of the overlapping \emph{and} disjoint communities users belong to.

Na\"ively the activity-based approach is motivated by the question of ``Which users in a network have similar activity profiles?".%, \emph{i.e.,} which user(s) cause groups of users to become active on the network.  %, \emph{i.e.}, which user(s) cause a user(s) to interact with the network. 
 With this question in mind, a community can then be thought of as those users who use (or do not use) the service at similar times. However, the measure we use is slightly more subtle that this. In particular, these communities can be thought of as groups of users who possess so-called ``predictive capacity" with one another, i.e., given a user $u$ and follower $f$ who are members of a community detected with this weighting, there exists a reduction in uncertainty about the activity profile of follower $f$ (tweeting or remaining silent), given the tweet history of user $u$.

@@

Finally, the interaction-based approach relies on the meta-data and text of messages to identify whom a user converses with on the social media service. On Twitter, we can use mentions (indicating a directed communication) and retweets (indicating endorsement of another user) to identify conversation. \alert{Wonder if the following sentence can be removed, to remove confusion between this and activity} Moreover, we can define a directed influence between two users by considering the attention paid to that user compared to all other users. This allows us to generate a network based on conversations and user interactions.
A few works have investigated this type of community. For example, \cite{conover2011political} considered both mention and retweet networks in isolation for a collection of users chosen for their political orientation. In~\cite{deitrick2013mutually}, the authors construct a dynamic network based on simple time-windowed counts of mentions and retweets, and use the evolution of this network to aid in community detection.

@@

Studying the intersections and lack-thereof of the communities derived from follower-followee, activity-based, topic-based, and interaction-based networks allow us to build a more complete picture of the \emph{implicit} networks present in online social media, as opposed to the \emph{explicit} social network indicated by structural links alone. In this paper, we explore the relation between these various possible networks and their corresponding communities. We begin by describing the methodologies used to generate the various types of networks, and infer their community structure. We then explore how the communities of users differ depending on the type of network used. Finally, we explore how communication patterns differ across and within the different community types. We conclude by illustrating that this multi-faceted question-oriented approach to community detection provides interesting insight into the intricate multi-faceted structure of several different users communities; information that would not be available if only a single round of community detection was performed. 
%\section{Related Work}

>>>>

%A large body of work exists on methods for automatic detection of communities within networks~\cite{newman2004fast,newman2004finding,rosvall2008maps,blondel2008fast,LancichinettiPlos}. See~\cite{fortunato2010community} for a recent review. 
A large body of work exists on methods for automatic detection of communities within networks, {\it e.g.,} the stochastic blockmodel~\cite{Holland1983} (and its recent generalization to weighted networks~\cite{Aicher26062014}), the Girvan-Newman algorithm~\cite{newman2004finding}, clique percolation~\cite{PalEtAl05}, Infomap~\cite{Rosvall08mapsof}, Louvain~\cite{blondel2008fast}, and the recently introduced OSLOM~\cite{LancichinettiPlos}.
All these methods begin with a given network, and then attempt to uncover structure present in the network, \emph{i.e.,} they are agnostic to how the network was constructed. As opposed to this agnostic analysis, we propose---and illustrate the importance of---a \emph{multi-faceted question-focused} approach. We believe that in order to understand all of the communities present in a network, it is important to take two crucial steps: 
\begin{enumerate}
\item Form a clear question about what \emph{type} of community we wish to consider---{\it e.g.,} are we looking for groups of friends? or for groups of people having a common interest?---and then perform community detection with that question in mind.
\item Ask more than one question about the users in a network. Do several rounds of community detection, each of which is aimed at revealing a new facet of the communities present in the network. 
\end{enumerate}

@@

This is not meant to be an exhaustive list, but rather a list of some of the more common types of communities observed in social networks. We propose looking at when and how communities motivated by these different questions overlap, and whether different approaches to asking the question, ``What community are you in?'' leads to different insights about a social network. For example, a user on Twitter might connect mostly with computational social scientists, utilize the service nearly every time a particular user or group of users is active, talk mostly about machine learning, and interact solely with close friends (who may or may not be computational social scientists). 
For this user, the answers to each question therefore correspond to more or less distinct communities. By contrast, a teenage user may only connect and interact with their friends, will most likely exhibit similar activity patterns dictated by the academic and extracurricular schedule of a student. For this user, the answers to each question map to the same community of users.
Thus, each question acts to highlight a particular `profile' of a user, and these profiles can help identify the ways in which a user interacts with and like other users in their online social network. 
We now consider in more detail how each of these types of questions motivates a question-specific community. Later in Section~\ref{sec:conc} we will show how asking \emph{all} of these questions---instead of focusing on only one---gives a much deeper multi-faceted view of the overlapping \emph{and} disjoint communities that users belong to.

Na\"ively the activity-based approach is motivated by the question of ``Which users in a network have similar activity profiles?"%, \emph{i.e.,} which user(s) cause groups of users to become active on the network.  %, \emph{i.e.}, which user(s) cause a user(s) to interact with the network. 
 With this question in mind, a community can then be thought of as those users who use (or do not use) the service at similar times. However, the measure we use is slightly more subtle that this. In particular, these communities can be thought of as groups of users who possess so-called ``predictive capacity" with one another, i.e., given a user $u$ and follower $f$ who are members of a community detected with this weighting, there exists a reduction in uncertainty about the activity profile of follower $f$ (tweeting or remaining silent), given the tweet history of user $u$.

@@

Finally, the interaction-based approach relies on the meta-data and text of messages to identify whom a user converses with on the social media service. On Twitter, we can use mentions (indicating a directed communication) and retweets (indicating endorsement of another user) to identify conversation. 
A few works have investigated this type of community. For example, \cite{conover2011political} considered both mention and retweet networks in isolation for a collection of users chosen for their political orientation. In~\cite{deitrick2013mutually}, the authors construct a dynamic network based on simple time-windowed counts of mentions and retweets, and use the evolution of this network to aid in community detection.

@@

Studying the intersections and lack-thereof of the communities derived from follower-followee, activity-based, topic-based, and interaction-based networks allow us to build a more complete picture of the \emph{implicit} networks present in online social media, as opposed to the \emph{explicit} social network indicated by follower links alone. In this paper, we explore the relation between these various possible networks and their corresponding communities. We begin by describing the methodologies used to generate the various types of networks, and infer their community structure. We then explore how the communities of users differ depending on the type of network used. Finally, we explore how communication patterns differ across and within the different community types. We conclude by illustrating that this multi-faceted question-oriented approach to community detection provides interesting insight into the intricate multi-faceted structure of several different users' communities; information that would not be available if only a single form of community detection was performed. 
%\section{Related Work}

### 2014-11-18 paper-postreview/methods.tex

The dataset for this study consisted of the tweets of 15,000 Twitter users over a 9 week period (from April 25th to June 25th 2011). The users are embedded in a network collected by performing a breadth-first expansion from a random seed user. In particular, once the seed user was chosen, the network was expanded to include his/her followers, but only included users considered to be `active' (i.e., users who tweeted at least once per day over the past one hundred days). Network collection continued in this fashion by considering the active followers of the active followers of the seed, and so on until 15,000 users were added to the network. The only biasing steps in this procedure are the selection of the seed node and the filtering out of inactive users. However, because all of the community types we are interested in involve some aspect of the interaction between users, each other, and their content, the procedure provides a relatively question-agnostic follower-followee network for use with the different question-based analyses.

Since our goal is to explore the functional communities of this network, we filter the network down to the subset of users which actively interact with each other (e.g., via retweets and mentions). We do this by measuring what we call (incoming/outgoing) information events. We define an outgoing information event for a given user $u$ as either a mention made by $u$ of another user in the network, or a retweet of one of $u$'s tweets by another user in the network. The logic for this definition is as follows: if $u$ mentions a user $v$ this can be thought of as $u$ directly sending information to $v$, and if $u$ is retweeted by $v$ then $v$ received information from $u$ and rebroadcast it to their followers. In either case there was information outgoing from $u$ which affected the network in some way. Analogously, we define the incoming information event for $u$ as either being mentioned by a different user in the network, or as retweeting another user in the network.
%function and information share is important We then filter down this starting %network to include only users that [[what??]]. To do this 
%Since one of the community types we want to explore is based on the explicit interactions ( between users, we filter the dataset in order to take into account only users who make use of these kinds of features in their tweets. [[Seeems out of place]]
%We define an event of outgoing information for a given user $u$ as either a mention made by $u$ of another user in the network, or a retweet by another user of one of $u$'s tweets. 
%[[Switched the readers perspective]]When we mention someone we are in fact sending him some information, and when we are retweeted it means that the person that retweeted us has received some information from us and is sharing it. 
With (incoming/outgoing) information events defined we filtered the network by eliminating all users with less than a total of 9 outgoing \emph{and} incoming information events, {\it i.e.,} less than one information event per type per week on average. 
We then further restricted our analysis to the strong giant connected component of the network built from the (incoming/outgoing) information filtered set of users. %and whose link represent a user-follower relationship.
In this study the link is directed from the user to the follower because this is the direction in which information flows. Thus, for a pair of users $u$ and $v$, an edge $a_{v \to u}$ in the structural network has weight 1 if user $u$ follows $v$, and 0 otherwise. The final network consists of 6,917 nodes and 1,481,131 edges.

>>>>

The dataset for this study consisted of the tweets of 15,000 Twitter users over a 9 week period (from April 25th to June 25th 2011). The users are embedded in a network collected by performing a breadth-first expansion from a random seed user. In particular, once the seed user was chosen, the network was expanded to include his/her followers, but only included users considered to be `active' (i.e., users who tweeted at least once per day over the past one hundred days). Network collection continued in this fashion by considering the active followers of the active followers of the seed, and so on until 15,000 users were added to the network. The only biasing steps in this procedure are the selection of the seed node and the filtering out of inactive users. However, because all of the community types we are interested in involve some aspect of the interaction between users, each other, and their content, this procedure provides a reasonable follower-followee network for use with the different question-based analyses.

Since our goal is to explore the functional communities of this network, we filter the network down to the subset of users which actively interact with each other (e.g., via retweets and mentions). We do this by measuring what we call (incoming/outgoing) information events. We define an outgoing information event for a given user $u$ as either a mention made by $u$ of another user in the network, or a retweet of one of $u$'s tweets by another user in the network. The logic for this definition is as follows: if $u$ mentions a user $v$ this can be thought of as $u$ directly sending information to $v$, and if $u$ is retweeted by $v$ then $v$ received information from $u$ and rebroadcast it to their followers. In either case there was information outgoing from $u$ which affected the network in some way. Analogously, we define the incoming information event for $u$ as either being mentioned by a different user in the network, or as retweeting another user in the network.
%function and information share is important We then filter down this starting %network to include only users that [[what??]]. To do this 
%Since one of the community types we want to explore is based on the explicit interactions ( between users, we filter the dataset in order to take into account only users who make use of these kinds of features in their tweets. [[Seeems out of place]]
%We define an event of outgoing information for a given user $u$ as either a mention made by $u$ of another user in the network, or a retweet by another user of one of $u$'s tweets. 
%[[Switched the readers perspective]]When we mention someone we are in fact sending him some information, and when we are retweeted it means that the person that retweeted us has received some information from us and is sharing it. 
With (incoming/outgoing) information events defined we filtered the network by eliminating all users with less than a total of 9 outgoing \emph{and} incoming information events, {\it i.e.,} less than one information event per type per week on average. 
We then further restricted our analysis to the strong giant connected component of the network built from the (incoming/outgoing) information filtered set of users. %and whose link represent a user-follower relationship.
In this study the link is directed from the user to the follower because this is the direction in which information flows. Thus, for a pair of users $u$ and $v$, an edge $a_{v \to u}$ in the follower-followee network has weight 1 if user $u$ follows $v$, and 0 otherwise. The final network consists of 6,917 nodes and 1,481,131 edges.

### 2014-11-18 paper-postreview/paper.tex
first round of edits

\title{Followers Are Not Enough: A Question-Oriented Approach to Community Detection in Online Social Networks}

@@

Community detection in online social networks is typically based on the analysis of the explicit connections between users, such as ``friends" on Facebook and ``followers" on Twitter. But online users often have hundreds or even thousands of such connections, and many of these connections do not correspond to real friendships or more generally to accounts that users interact with. We claim that community detection in online social networks should be question-oriented and rely on additional information beyond the simple structure of the network. The concept of `community' is very general, and different questions such as ``whom do we interact with?" and ``with whom do we share similar interests?" can lead to the discovery of different social groups. In this paper we focus on three types of communities beyond structural communities: activity-based, topic-based, and interaction-based. We analyze a Twitter dataset using three different weightings of the structural network meant to highlight these three community types, and then infer the communities associated with these weightings. We show that the communities obtained in the three weighted cases are highly different from each other, and from the communities obtained by considering only the unweighted structural network. Our results confirm that asking a precise question is an unavoidable first step in community detection in online social networks, and that different questions can lead to different insights about the network under study.
%\end{quote}
\end{abstract}

>>>>

\title{Followers Are Not Enough: A Multi-faceted Approach to Community Detection in Online Social Networks}

@@

Community detection in online social networks is typically based on the analysis of the explicit connections between users, such as ``friends" on Facebook and ``followers" on Twitter. But online users often have hundreds or even thousands of such connections, many of which connect them with news outlets, governmental organizations, etc., in addition to their friends and associates. Because of the multi-faceted nature of these networks, we claim that community detection in online social networks should be question-oriented and rely on additional information beyond the simple structure of the network. The concept of `community' is very general, and different questions such as ``whom do we interact with?" and ``with whom do we share similar interests?" can lead to the discovery of different social groups. In this paper we focus on three types of communities beyond structural communities: activity-based, topic-based, and interaction-based. We analyze a Twitter dataset using three different weightings of the structural network meant to highlight these three community types, and then infer the communities associated with these weightings. We show that the communities obtained in the three weighted cases are highly different from each other, and from the communities obtained by considering only the unweighted structural network. Our results confirm that asking a precise question is an unavoidable first step in community detection in online social networks, and that different questions can lead to different insights about the network under study.
%\end{quote}
\end{abstract}
